---
title: "Descriptive Statistics"
author: "Ryan"
date: "September 12, 2015"
output:
  pdf_document: 
    highlight: espresso
  html_document:
    highlight: espresso
    smart: no
    theme: readable
---

# Descriptive Statistics
## You now know vectors...
## Lets generate some data in vector form
```{r}
set.seed(123)
ranks <- as.factor(sample(10,40,replace = T ))
ranks
set.seed(234)
heights <- rnorm(40, 171, 5)
heights
```


# Descriptive Statistics for Categorical data
## Frequency Distribution Table
```{r}
ft <- table(ranks)
ft
```

## Relative Frequency Distribution
```{r}
ft/length(ranks)
```

## Percentage Relative Frequency Distribution
```{r}
ft/length(ranks)*100
```

## Bar Chart
```{r}
barplot(ft, space = 0, border = 0)
```

## Pie Chart
```{r}
pie(table(ranks), border = 0)
```

# Descriptive Statistics for Numerical data
## Frequency Distribution Table
```{r}
b <- seq(154, 181, 3)
heights.cut <- cut(heights, breaks = b, right = F)
cft <- table(heights.cut)
cft
```

## Relative Frequency Distribution
```{r}
cft/length(heights)
```

## Percentage Relative Frequency Distribution
```{r}
cft/length(heights) * 100
```

## Histogram
```{r}
barplot(cft, border = 0, space = 0)
```

```{r}
hist(heights)
```

```{r}
hist(heights, breaks = b, right = F)
```

## Measure of Centrality: Mean and Median
```{r}
mean(heights)
median(heights)
```

## Quantiles
```{r}
quantile(heights, probs = 0.25, type = 6)
quantile(heights, probs = 0.5, type = 6)
quantile(heights, probs = 0.75, type = 6)
quantile(heights, probs = 0.95, type = 6)
```

## Five Number Summmary
```{r}
summary(heights)
```

## Boxplot
```{r}
boxplot(heights, horizontal = T, pch = 8)
```

## Measure of Dispersion : Variance and Standard Deviation
```{r}
var(heights)
sd(heights)
```

# Descriptive Statistics for Linear Relationship
## Scatter plot
```{r}
set.seed(231)
weights <- heights/3+rnorm(40,0,0.5)
plot(weights, heights, pch = 19)
lines(x = rep(mean(weights),2), y = c(min(heights),max(heights)))
lines(y = rep(mean(heights),2), x = c(min(weights),max(weights)))
```

## Covariance and Correlation Coefficient
```{r}
cov(heights, weights)
cor(heights, weights)
```

