---
title: "Data Table,Dplyr,Manipulate,Shiny"
author: "Ryan Zhang"
date: "November 5, 2015"
fontsize: 8pt
output:
  beamer_presentation:
    colortheme: whale
    fig_height: 4
    fig_width: 8
    fonttheme: serif
    highlight: espresso
    theme: Warsaw
---

```{r, echo = F}
setwd("D:/GithubRepos/Think_R/2015 R Study Group/Nov 6th")
```

## Data Table
```{r, warning=F, message= F}
library(data.table)
```
+ Fast aggregation of large data
+ Fast add/modify/delete of columns by group using no copies at all
+ Fast file reader (fread). 

## Data table
+ Example data
+ Thanks, thanks, thanks Monica !
```{r}
emp <- fread("emp_t.csv")
dept <- fread("dep_t.csv")
product <- fread("product_t.csv")
price <- fread("price_t.csv")
order <- fread("order_t.csv")
orderline <- fread("orderline_t.csv")
customer <- fread("customer_t.csv")
```

## Data table
+ If a data.table/data.frame object is been called by a function that only knows data.frame, it will be passed as a data.frame
```{r}
class(emp)
```
+ Basic data.table syntax
```
DT [i,
    .(j1 = col1, j2 = col2, j3 = func1(col3)),
    by = .(col4, col5)]
```

## Data Table
+ Provide a shortcut `.N` for number of rows in data table
```{r}
emp[1:(nrow(emp)/4)]
emp[1:(.N/4)]
```

## Data Table
+ Subset by row, similar to the following SQL query:    
```
SELECT * 
FROM emp
WHERE MGR = 7698;
```   
```{r}
emp[MGR == 7698]
```

## Data Table
+ Subset by row, and specify columns wanted:    
```
SELECT EMPNO, EFNAME, ELNAME, EMP_JOB 
FROM emp
WHERE MGR = 7698;
```   
```{r}
emp[MGR == 7698,
    .(EMPNO, EFNAME, ELNAME, EMP_JOB)]
```

## Data Table
+ Provide a method to apply function on columns
```{r}
c(min(emp$SAL), mean(emp$SAL))
emp[, 
    .(MinSalary = min(SAL), AvgSalary = mean(SAL))]
```

## Data Table
+ More than math function can be used on columns in data table
```{r}
emp[, 
    .(plot(SAL, pch = 19))]
```

## Data Table
+ Group by in data table
+ Getting more similar to SQL now...    
```
SELECT DEPTNO, AVG(SAL) AS "AvgSalary"
FROM emp
GROUP BY DEPTNO;
```
```{r}
emp[, 
    .(AvgSalary = mean(SAL)),
    by = .(DEPTNO)]
```

## Data Table
+ More powerful group by than SQL
+ Compare the mean Salary of SALESMAN and NON-SALESMAN
+ Need two SQL queries or sub queries to get the result
```{r}
emp[, 
    .(AvgSalary = mean(SAL)),
    by = .(EMP_JOB == "SALESMAN")]
```

## Data Table
+ Compare the mean Salary of SALESMAN and NON-SALESMAN within DEPTNO == 20
+ Think about how to do it in SQL for a minute...
```{r}
emp[DEPTNO == 20, 
    .(AvgSalary = mean(SAL)),
    by = .(EMP_JOB == "SALESMAN")]
```

## Data Table
+ Find the lowest paid employees name
+ You need nested Select in SQL, but much simpler in data.table
```
SELECT EFNAME, ELNAME 
FROM emp 
WHERE SAL = (SELECT MIN(SAL) FROM emp);
```
```{r}
emp[SAL == min(SAL), 
    .(EFNAME, ELNAME)]
```

## Data Table
+ Remember: what's been returned is a data table
+ Find all orders that have more than 3 orderlines on them. List out the order id
```
SELECT orderid
FROM orderline
GROUP BY orderid
HAVING count(*) > 3;
```
```{r}
orderline[,
          .(number = .N),
          by = .(ORDERID)][number >3, 
                           .(ORDERID)]
```

## sqldf
+ sqldf allow you use SQLite syntax to query dataframes
```{r message=F, warning=F}
library(sqldf)
sqldf("SELECT EFNAME, EMIDDLE_INIT, ELNAME, emp.DEPTNO
       FROM dept, emp
       WHERE dept.DEPTNO = emp.DEPTNO 
	        AND SAL > (SELECT MAX(SAL)
			           FROM dept, emp
       		           WHERE dept.DEPTNO = emp.DEPTNO AND emp.DEPTNO = 20);")
```

## R can work ontop of many DBMS
```{r, eval = F}
install.packages("RPostgreSQL")
install.packages("RMySQL")
install.packages("RMongo")
install.packages("RODBC")
install.packages("RSQLite")
```
+ They said there is also a `ROracle` library, but I have trouble installing it.  

## dplyr
+ They said, you can join tables using data.table, which I never figure out how to do....
+ Maybe you can find out and teach me how to do it.
+ If I need to join tables, I will use dplyr.
+ So... Let's talk about dplyr now.

## dplyr
```{r, message = F, warning = F}
library(dplyr)
```
+ Fast data bla bla bla
+ Can work on top of database
+ Very descriptive syntax, like SQL 
    - "tell me what you want"
    - "Not necessary how I should find it for you"

## dplyr
+ Verbs
    1. filter()
    2. slice()
    3. arrange()
    4. select()
    5. distinct()
    6. transform() 
    7. summarise()
    8. group_by()

## dplyr
+ filter() allows you to select a subset of rows by a set of conditions. 
```
SELECT PRODUCTFINISH, PRODUCTSTANDARDPRICE
FROM product
WHERE PRODUCTFINISH = "Cherry" AND PRODUCTSTANDARDPRICE > 200;
```
```{r}
filter(product, PRODUCTFINISH == "Cherry", PRODUCTSTANDARDPRICE > 200)
```

## dplyr
+ slice() allows you the select a subset of rows by positions:
+ How to do this in SQL?
+ LIMIT and OFFSET are supported by PostgreSQL, but not Oracle? (If my memory is correct.)  
```{r}
slice(orderline, 1:6)
```

## dplyr
+ arrange() reorders rows according to values in specified columns. 
+ Similar to ORDER BY in SQL. 
```
SELECT *
FROM product
ORDER BY PRODUCTFINISH, PRODUCTSTANDARDPRICE;
```
```{r}
arrange(product, PRODUCTFINISH, PRODUCTSTANDARDPRICE)
```

## dplyr
+ Use desc() to order a column in descending order.
```{r}
arrange(product, PRODUCTFINISH, desc(PRODUCTSTANDARDPRICE))
```

## dplyr
+ select() allows you to select columns:
```
SELECT STDPRICE, MINPRICE
FROM price;
```
```{r}
select(price, STDPRICE, MINPRICE)
```

## dplyr
+ distinct() to find unique values in a table.
+ Remind you of the DISTINCT in SQL?
```
SELECT DISTINCT PRODUCTFINISH
FROM product;
```
```{r}
distinct(select(product, PRODUCTFINISH))
```

## dplyr
+ Add new columns with mutate(), you can directly work on the columns you just created.
+ Think about how to do this in SQL. 
```{r, echo = 2:4}
product2 <- tbl_df(product)
mutate(product2, 
       DISCOUNT = ifelse(PRODUCTFINISH == "Leather", 0.85, 0.7),
       DISCOUNTEDPRICE = DISCOUNT * PRODUCTSTANDARDPRICE)
```

## dplyr
+ use transmute() is only include the columns created
```{r}
transmute(product2, 
            DISCOUNT = ifelse(PRODUCTFINISH == "Leather", 0.85, 0.7),
            DISCOUNTEDPRICE = DISCOUNT * PRODUCTSTANDARDPRICE)
```

## dplyr
+ use summarise() to do aggregation
```
SELECT AVG(PRODUCTSTANDARDPRICE)
FROM product;
```
```{r}
summarise(product,
  meanPrice = mean(PRODUCTSTANDARDPRICE))
```

## dplyr
+ use group_by() to do grouping.   
```
SELECT AVG(PRODUCTSTANDARDPRICE)
FROM product
GROUP BY PRODUCTFINISH;
```
```{r}
product_by_FINISH <- group_by(product, PRODUCTFINISH)
summarise(product_by_FINISH,
          meanPrice = mean(PRODUCTSTANDARDPRICE))
```

## dplyr
+ Chaining. 
```
SELECT AVG(PRODUCTSTANDARDPRICE) AS "meanPrice"
FROM product
GROUP BY PRODUCTFINISH
HAVING AVG(PRODUCTSTANDARDPRICE) > 300
ORDER BY AVG(PRODUCTSTANDARDPRICE) DESC;
```
```{r}
arrange(
    filter(
        summarise(group_by(product, 
                           PRODUCTFINISH),
                  meanPrice = mean(PRODUCTSTANDARDPRICE)),
        meanPrice > 300), 
    desc(meanPrice))
```

## dplyr
+ Use `%>%` this strange symbol.. To make code more readable. 
```{r}
product %>%
    group_by(PRODUCTFINISH) %>%
    summarise(meanPrice = mean(PRODUCTSTANDARDPRICE)) %>%
    filter(meanPrice > 300) %>%
    arrange(desc(meanPrice))
```

## dplyr
+ Lots of join operations as well...
```{r, echo = 2}
emp$DEPTNO <- as.integer(emp$DEPTNO)
inner_join(emp, dept)
```

## dplyr
+ Find all Products purchased by at least 4 customers.
```
SELECT PRODUCTID
FROM order, orderline 
WHERE order.ORDERID = orderline.ORDERID
GROUP BY PRODUCTID, CUSTOMERID
HAVING COUNT(*) >= 4;
```
```{r}
inner_join(order, orderline, by = "ORDERID") %>%
    group_by(PRODUCTID, CUSTOMERID) %>%
    summarise(countCustomer = n()) %>%
    filter(countCustomer >= 4) %>%
    select(PRODUCTID)
```

## dplyr
+ Find all employees who make more than their manager
```
SELECT EFNAME, EMIDDLE_INIT, ELNAME
FROM emp employee, emp manager
WHERE employee.mgr = manager.empno 
    AND employee.SAL > manager.SAL;
```
+ I don't know how to do self join using dplyr
+ Maybe you can dig into it and teach me...

## manipulate
+ Interactive plotting functions for use within RStudio.
+ The manipulate function accepts a plotting expression and a set of controls (e.g. slider, picker, checkbox, or button) which are used to dynamically change values within the expression. 
```{r message = F, warning = F, eval=F}
library(manipulate);library(ggplot2)
manipulate(
qplot(x = diamonds[,variable], geom = geomtype, fill = diamonds[, colorby])+
    xlim(c(xmin, xmax)),
variable = picker("carat" = "carat", "depth" = "depth", "table" = "table", 
                  "price" = "price", "x" = "x", "y" = "y","z" = "z"),
geomtype = picker("histogram" = "histogram", "freqpoly" = "freqpoly"),
colorby = picker("cut" = "cut", "color" = "color", "clarity" = "clarity"),
xmin = slider(0,300), xmax = slider(5.2,19000))
```

## shinyapps
+ Shinyapps.io is a platform as a service (PaaS) for hosting Shiny web apps (applications). 
+ ui.R
+ server.R
```{r, eval=F}
library(shiny)
runApp()
```

## shinyapps
+ ui.R
```{r, eval =F}
library(shiny)
shinyUI(pageWithSidebar(
  headerPanel("Standard Normal Density"),# Application title
  sidebarPanel(# Sidebar with a slider input for 
    p("computes probability for a standard normal distribution."),
    p("choose a score by sliding the slide, and choose tail using the dropdown."),
    h3('Please choose score'),
    sliderInput("score", "It could only be between -4 and 4:", 
                min = -4, max = 4, value = 0, step = 0.01),
    h3('Please choose lower tail or upper tail'),
    selectInput("lo", " ", choices = c("lower", "upper"), selected = "lower")
    ),
  mainPanel(# show a normal distribution with score indicated as a red vertical line
    plotOutput("distPlot"),
    h3("The associated probability is"),
    verbatimTextOutput("pvalue"))))
```

## shinyapps
+ server.R
```{r eval= F}
library(shiny)
shinyServer(
  function(input,output){
    output$distPlot <- renderPlot({# generate an rnorm distribution and plot it
      x <- seq(-4,4,by = 0.01)
      y <- dnorm(x,mean=mean(x),sd=sd(x)) 
      xl <- rep(input$score,301)
      yl <- seq(0,0.3,by = 0.001)
      plot(x, y, type = "l",col="black", lwd=2,ylab = "Pr(X=x)")
      lines(xl,yl,col="red",lwd=2)
    })
    lower <- reactive({switch(input$lo, "lower" = T, "upper" = F)})
    output$pvalue <- renderPrint({pnorm(input$score, 0,1, lower.tail = lower())})
    })
```



