---
title: "ST625 Chapter 3 Example using R"
author: "Ryan"
date: "September 7, 2015"
output:
  html_document:
    highlight: espresso
  pdf_document:
    highlight: espresso
---

# 3.9 A complete example
## Step1. Model is $y = \beta_0 + \beta_1x + \epsilon$
```{r}
formula = formula('damageY~distanceX')
```

## Step2. get the data
fire damage data
```{r}
fire <- cbind.data.frame(
        distanceX = c(3.4,1.8,4.6,2.3,3.1,5.5,.7,3.0,
                      2.6,4.3,2.1,1.1,6.1,4.8,3.8),
        damageY = c(26.2,17.8,31.3,23.1,27.5,36.0,14.1,
                    22.3,19.6,31.3,24.0,17.3,43.2,36.4,26.1))
```

## Step3 fit the model using the data
model fitting using lm function
```{r}
fireModel <- lm(formula, data = fire)
fireModel$coefficients
```

model fitting using formula
```{r}
SSxx = sum((fire$distanceX-mean(fire$distanceX))^2)
SSxy = sum((fire$distanceX-mean(fire$distanceX))*(fire$damageY-mean(fire$damageY)))
b1 = SSxy/SSxx
b0 = mean(fire$damageY) - b1 * mean(fire$distanceX)
b0;b1
```

model fitting using linear algebra
```{r}
X = as.matrix(cbind(rep(1,nrow(fire)),fire$distanceX))
Y = as.matrix(fire$damageY)
beta = solve(t(X) %*% X) %*% (t(X) %*% Y)
beta
```


scatter plot with the regression line
```{r}
plot(x = fire$distanceX, y = fire$damageY, pch = 19)
abline(fireModel, lwd = 2)
```

## Step4. Check assumptions on random error
### 1 $\mathbb{E}(\epsilon)=0$
```{r}
mean(fireModel$residuals)
```

### 2 Var($\epsilon$) is constant 
```{r}
MSE = sum(fireModel$residuals^2)/fireModel$df.residual
MSE
RMSE = sqrt(MSE)
plot(fire$distanceX, fireModel$residuals, pch = 19, ylim = c(-5,5))
abline(0,0)
abline(2.316,0, col = 'red')
abline(-2.316,0, col = 'red')
abline(2 * 2.316,0, col = 'blue')
abline(2 * -2.316,0, col = 'blue')
```

### 3 $\epsilon$ normally distributed
using histogram
```{r}
hist(fireModel$residuals, border = F, col = 'black')
```

using qq plot
```{r}
qqnorm(fireModel$residuals, pch = 19)
abline(0,1)
```

coefficient of variation
```{r}
100*RMSE/mean(fire$damageY)
```

Summary information
```{r}
summary(fireModel)
```

## Step5. Assess model adequacy
### a Test of model utility
```{r}
SE = RMSE/sqrt(SSxx)
t = b1/SE
2*pt(-1*abs(t), df = fireModel$df.residual)
```

### b Confidence interval for slope
using the R function
```{r}
confint(fireModel)
```

hand calculation
```{r}
b1 + qt(0.025, df = fireModel$df.residual) * SE * c(1,-1)
```

### c Numerical descriptive measures of model adequacy
R Squared calculated by hand
```{r}
SSE = sum(fireModel$residuals^2)
SSyy = sum((fire$damageY-mean(fire$damageY))^2)
RSquare = 1- SSE/SSyy
RSquare
summary(fireModel)
```

Coefficient of correlation
by hand and using function
```{r}
cor(fire$damageY, fire$distanceX)
SSxy/sqrt(SSxx*SSyy)
```

### Step6. Prediction
ignore the tideous calculation
```{r}
linspace = seq(0,7,0.01)
confInterval <- predict(fireModel,
                        newdata = data.frame(distanceX = linspace),
                        interval = "confidence")
predInterval <- predict(fireModel,
                        newdata = data.frame(distanceX = linspace),
                        interval = "prediction")

plot(x = fire$distanceX,
     y = fire$damageY,
     pch = 19,
     ylim=c(0,50),
     col = adjustcolor('black', alpha.f =  0.8))
abline(fireModel, lwd = 2)
points(x= linspace, y = confInterval[,2], type = 'l', col = 'red')
points(x= linspace, y = confInterval[,3], type = 'l', col = 'red')
points(x= linspace, y = predInterval[,2], type = 'l', col = 'blue')
points(x= linspace, y = predInterval[,3], type = 'l', col = 'blue')
```

Make perdiction on perticular value
```{r}
predict(fireModel,
        newdata = data.frame(distanceX = 3.5),
        interval = "prediction")
```