---
title: "Kaggle Titanic Example"
author: "Ryan"
date: "September 7, 2015"
output:
  html_document:
    highlight: espresso
    theme: cosmo
  pdf_document:
    highlight: espresso
    theme: cosmo
---

## load data
```{r chunck1}
train <- read.csv(file.choose(), 
                  header = T,
                  stringsAsFactors = F)
```

## look at data
View using spreadsheet like window
```{r chunck2}
View(train)
```

Print 5 lines in console
```{r chunck3}
head(train)
```

view the structure
```{r chunck4}
str(train)
```

## prepare for model fitting
Covert some variables into categorical(factor) variable
```{r chunck5}
train$Survived <- factor(train$Survived)
train$Pclass <- factor(train$Pclass)
train$Sex <- factor(train$Sex)
```

## Modeling
model formula
```{r chunck6}
formula = formula("Survived~Pclass+Sex+Age")
```

logistic regression model
$$y = \frac{1}{1+e^{-\beta_0-\beta_1x}}$$

```{r chunck7}
model <- glm(formula, data = train, family = "binomial")
```

in sample prediction accuracy using the model
look at the contingency table
```{r chunck8}
conTable <- table(predict(model, train) > 0.5, train$Survived)
conTable
(117+392)/sum(conTable)
```

## Make prediction on test data

```{r chunck9}
test <- read.csv(file.choose(), 
                 header = T, 
                 stringsAsFactors = F)
str(test)
test$Pclass <- factor(test$Pclass)
test$Sex <- factor(test$Sex)
testpredict <- predict(model, test)
testpredict <- ifelse(is.na(testpredict),
                      runif(1,0,1),
                      testpredict)
prediction <- ifelse(testpredict > 0.5,
                     1,
                     0)
submit <- cbind.data.frame(PassengerId = test$PassengerId,
                           Survived = prediction)
write.csv(file = "simpleLogisticRegression.csv",
          submit, 
          row.names = F)
```

## Classification tree model
```{r}
#install.packages("rpart")
library(rpart)
rpartModel <- rpart(formula, train)
rpartPred <- predict(rpartModel, test, type ="class")
submit <- cbind.data.frame(PassengerId = test$PassengerId,
                           Survived = rpartPred)
write.csv(file = "rpartDefault.csv",
          submit, 
          row.names = F)
```

