---
title: "K-Means"
author: "Ryan"
date: "October 4, 2015"
output: html_document
---

## Plotting setting
```{r}
library(ggplot2)
library(ggthemes)
theme_set(theme_minimal())
```

## Some example data
What should the k be ?  
```{r}
data <- data.frame(x = c(1,0,2,5,4,5,1,0,2,6,6,4,3,3,4,4,5,5),
                   y = c(2,0,1,3,1,0,-1,1,0,2,1,2,6,7,8,5,6,7))
ggplot(aes(x = x, y = y), data = data) + 
    geom_point(size = 5) + 
    xlim(-2,8) + 
    ylim(-2,8)
```

## Distance Measures
I am using euclidean distance here    
But there are many other options as well   

define your own distance Function   
```{r}
eucDistance <- function(p1,p2){
    return(sqrt(sum((p1-p2)^2)))
}

eucDistance(data[1,],data[2,])
eucDistance(data[3,],data[4,])
eucDistance(data[1,],c(0,0))
```

remember the loop construct?
```{r}
n = nrow(data)
for (i in 1:n) {
    print(eucDistance(data[i,],c(0,0)))
}
```

## Centroids  
centroids are just points   
however their positions may be updated in each iteration    
Initialize three centorids    
```{r}
centroids <- data.frame(x = c(0,7,4),
                        y = c(3,2,5))
ggplot(aes(x = x, y = y), data = data) + 
    geom_point(size = 5) + 
    xlim(-2,8) + 
    ylim(-2,8) +
    geom_point(aes(x = x, y = y), data = centroids, size = 8, color = 'blue', alpha = 0.5)
```

## Assign points the centroids    
Obivously assign points to the centroid that is closest to it    
Let's see to which centroid point (0,0) should be assigned to?    
```{r}
for (cid in 1:3){ # we got three centroids
    print(eucDistance(c(0,0), centroids[cid,]))
}
```

So we should assign point (0,0) to centroid 1  

```{r}
distances <- vector()
for (cid in 1:3){
    distances <- c(distances, eucDistance(data[1,], centroids[cid,]))
}
distances
which.min(distances)
```

Using nested loops to figure out the assignment for all points   
```{r}
assignment <- vector()
for (pid in 1:n){ # we got n points
    distances <- vector()
    for (cid in 1:3){ # we got 3 centroids
        distances <- c(distances, eucDistance(data[pid,], centroids[cid,]))
    }
    assignment <- c(assignment, which.min(distances))
}
assignment
data$assignment <- assignment
data
```

## update the position of the centroids  
Update to what position ?    
The center of the points assigned to it...    
That is why they are called centroids, aren't they ?   :)
also using loop    
```{r}
for (cid in 1:3){
    centroids[cid,] = apply(data[data$assignment == cid,], 2, mean)[1:2]
}
centroids
ggplot(aes(x = x, y = y), data = data) + 
    geom_point(size = 5) + 
    xlim(-2,8) + 
    ylim(-2,8) +
    geom_point(aes(x = x, y = y), data = centroids, size = 8, color = 'blue', alpha = 0.5)
```

## Converge
Done... if your run the code again, nothing will change    

## Make it more general than 3 centroids
```{r}
myKMeans <- function(data, centroids, maxIter = 999){
    n = nrow(data)    
    m = ncol(data)
    k = nrow(centroids)
    data$assignment <- k+1
    p <- ggplot(aes(x = x, y = y), data = data) + 
             geom_point(size = 5, color = data$assignment) + 
             xlim(-2,8) + 
             ylim(-2,8) +
             geom_point(aes(x = x, y = y), data = centroids, size = 25, color = 
                            as.numeric(row.names(centroids)), alpha = 0.5) + 
             geom_point(aes(x = x, y = y), data = centroids, size = 8, color = 'blue')
    plot(p)
    for (i in 1:maxIter){
        print(paste("iteration",i))
        assignment <- vector()
        oldAssignment <- data$assignment
        for (pid in 1:n){ # we got n points
            distances <- vector()
            for (cid in 1:k){ # we got 3 centroids
                distances <- c(distances, eucDistance(data[pid,1:m], centroids[cid,1:m]))
            }
            assignment <- c(assignment, which.min(distances))
        }
        print(assignment)
        data$assignment <- assignment
        for (cid in 1:k){
            centroids[cid,] = apply(data[data$assignment == cid,], 2, mean)[1:2]
        }
        print(centroids)
        if (sum(oldAssignment == data$assignment)== n) break   
        p <- ggplot(aes(x = x, y = y), data = data) + 
            geom_point(size = 5, color = data$assignment) + 
            xlim(-2,8) + 
            ylim(-2,8) +
            geom_point(aes(x = x, y = y), data = centroids, size = 25, color = 
                           as.numeric(row.names(centroids)), alpha = 0.5) + 
            geom_point(aes(x = x, y = y), data = centroids, size = 8, color = 'blue') 
        plot(p)
    
    }
    print("===============DONE===================")
    return(data)
}
rm(assignment)
```

See some example run of our k-means
```{r}
data <- data.frame(x = c(1,0,2,5,4,5,1,0,2,6,6,4,3,3,4,4,5,5),
                   y = c(2,0,1,3,1,0,-1,1,0,2,1,2,6,7,8,5,6,7))
centroids <- data.frame(x = c(1),
                        y = c(1))
cdata <- myKMeans(data, centroids)
```


```{r}
data <- data.frame(x = c(1,0,2,5,4,5,1,0,2,6,6,4,3,3,4,4,5,5),
                   y = c(2,0,1,3,1,0,-1,1,0,2,1,2,6,7,8,5,6,7))
centroids <- data.frame(x = c(0,7),
                        y = c(3,2))
cdata <- myKMeans(data, centroids)
```

```{r}
data <- data.frame(x = c(1,0,2,5,4,5,1,0,2,6,6,4,3,3,4,4,5,5),
                   y = c(2,0,1,3,1,0,-1,1,0,2,1,2,6,7,8,5,6,7))
centroids <- data.frame(x = c(2,5),
                        y = c(2,5))
cdata <- myKMeans(data, centroids)
```

```{r}
data <- data.frame(x = c(1,0,2,5,4,5,1,0,2,6,6,4,3,3,4,4,5,5),
                   y = c(2,0,1,3,1,0,-1,1,0,2,1,2,6,7,8,5,6,7))
centroids <- data.frame(x = c(7,7),
                        y = c(1,2))
cdata <- myKMeans(data, centroids)
```

```{r}
data <- data.frame(x = c(1,0,2,5,4,5,1,0,2,6,6,4,3,3,4,4,5,5),
                   y = c(2,0,1,3,1,0,-1,1,0,2,1,2,6,7,8,5,6,7))
centroids <- data.frame(x = c(7,7,4,3,1),
                        y = c(1,2,5,6,2))
cdata <- myKMeans(data, centroids)
```

```{r}
data <- data.frame(x = c(1,0,2,5,4,5,1,0,2,6,6,4,3,3,4,4,5,5),
                   y = c(2,0,1,3,1,0,-1,1,0,2,1,2,6,7,8,5,6,7))
centroids <- data.frame(x = c(1,2,3,4,5,6),
                        y = c(1,2,3,4,5,6))
cdata <- myKMeans(data, centroids)
```



## Key Lessons
1. K is important   
2. Initial position of centroids also matters a lot   

